---
title: "Section 7.2"
author: "Patrick Burke"
date: "2/16/2022"
output: html_document
---

```{r}
library(tidyverse)
library(ggplot2)
library(mapproj)
library(maps)
hpi_df <- read_csv("data/hpi-tidy.csv")
```

#Exercise 1
```{r}
map("world", projection = "mercator", wrap = TRUE) 

worldmap <- map_data("world")

hpi_world <- left_join(x = worldmap, y = hpi_df, 
                       by = c("region" = "Country"))

hpi_world <- hpi_world %>%
  filter(!is.na(hpi_world$LifeExpectancy))
```

```{r}
map1 <- ggplot(data = hpi_world, aes(x = long, y = lat, group=group)) +
  geom_polygon(aes(fill = HappyLifeYears), color = "black")
map1
```

#Exercise 2
```{r}
hpi_world2 <- left_join(x = worldmap, y = hpi_df, 
                       by = c("region" = "Country"))

map2 <- ggplot(data = hpi_world2, aes(x = long, y = lat, group=group)) +
  geom_polygon(aes(fill = HappyLifeYears), color = "black")

map2

```


#Exercise 3

I did not have to deal with this until exercise 5, so I you can see me fix that map down there.


#Exercise 4

I think this is a very interesting phenomenon. The fact that Africa is 550 times the size of Greenland, but on a map they appear as near equivalents is shocking.The same is visible with the United States and South America.


#Exercise 5
```{r}
map3 <- ggplot(data = hpi_world2, aes(x = long, y = lat, group=group)) +
  geom_polygon(aes(fill = HappyLifeYears), color = "black") +
  coord_map(projection = "globular")
map3
```

```{r}
map4 <- ggplot(data = hpi_world2, aes(x = long, y = lat, group=group)) +
  geom_polygon(aes(fill = HappyLifeYears), color = "black") +
  coord_map(projection = "gilbert")
map4
```
Now I got the two horizontal lines!


```{r}
map5 <- ggplot(data = hpi_world2, aes(x = long, y = lat, group=group)) +
  geom_polygon(aes(fill = HappyLifeYears), color = "black") +
  coord_map("gilbert", xlim=c(-180,180))
map5
```
And, fixed. It is interesting with the Gilbert projection, Africa and Greenland look closer in size than in the Globular projection. I believe the Globular projection is most likely more accurate, as I can see that if Greenland were to fit in Africa 550 times, it would have to be seen as is in the Globular projection.
