---
title: "Take Home Quiz - 5"
author: "Patrick Burke"
date: "2/22/2022"
output: html_document
---

```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(readxl)
stat113_df <- read_xlsx("data/stat113_survey.xlsx")
```

# Exercise 1

```{r}
stat113_mean <- stat113_df %>%
  filter(Facebook != "NA")

stat113_mean <- stat113_mean %>%
  mutate(Facebook = as.integer(Facebook))

stat113_mean <- stat113_mean %>%
  group_by(time_both)%>%
  summarise(friendmean = mean(Facebook))

stat113_df1 <- stat113_df %>%
  filter(Facebook != "NA") %>%
  mutate(Facebook = as.integer(Facebook))


stat113_df1$time_both <- ordered(stat113_df1$time_both, 
                                 levels = c("S13", "F13", "S14", "F14", "S15", 
                                            "F15", "S16", "F16", "S17", "F17", 
                                            "S18", "F18", "S19", "F19", "S20", 
                                            "F20", "S21", "F21"))
```
Here I set up my data sets to have an average amount of Facebook friends for the graph.

```{r}
ggplot(data = stat113_df1, aes(x = time_both, y = Facebook)) +
  geom_point(alpha = .3, color = "darkgray") +
  coord_flip() +
  geom_point(data = stat113_mean, 
             mapping = aes(x = time_both, y = friendmean, 
                           color = "green")) +
  labs(x = "Semester", y = "Number of Facebook Friends", color = "Semester Average")



```

Here is a graph that shows the variability of Facebook friends, as well as the average amount of friends per semester. You can see in this graph that the average Stat-113 student's Facebook friends have shrunk over recent years.

# Exercise 2

```{r}
stat113_df <- stat113_df %>%
  mutate(Facebook = as.integer(Facebook))

stat113_df2 <- stat113_df %>%
  mutate(FacebookCount = Facebook) %>%
  mutate(FacebookCount = as.integer(FacebookCount))

stat113_df2 <- stat113_df2 %>%
  mutate(FacebookCount = ifelse(Facebook > 0, 
                                TRUE, FALSE)) %>%
  mutate(FacebookCount = ifelse(is.na(FacebookCount), FALSE, TRUE))

stat113_count <- stat113_df2 %>%
  group_by(time_both) %>%
  mutate(FacebookProp = sum(FacebookCount == TRUE) / 5322) %>%
  filter(!is.na(time_both))

stat113_count$time_both <- ordered(stat113_count$time_both, 
                                 levels = c("S13", "F13", "S14", "F14", "S15", 
                                            "F15", "S16", "F16", "S17", "F17", 
                                            "S18", "F18", "S19", "F19", "S20", 
                                            "F20", "S21", "F21"))
```

```{r}
ggplot(stat113_count, aes(x = time_both, y = FacebookProp, color = time_semester)) +
  geom_col() +
  labs(x = "Semester", y = "Proportion of Facebook Users", color = "Which Semester")
```

Here is my final graph. This graph shows us the proportion of Facebook users among STAT-113 students throughout the years. We can see, as we saw in the previous graph of averages, that Facebook is past its peak; however, we can see there has been a larger proportion of Facebook users than before in the Fall of 2021.

